Sharing statistics for SPARQL Federation optimization, with emphasis on benchmark quality

### 2012-01-20 eswc-2012/phdsymp.tex
review paper
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@106 43ef1190-90fd-4dd7-a769-d0121f02b80e

I take the technology state of the art to be represented by the
current basic SPARQL 1.1 Federated Query Working
Draft\footnote{http://www.w3.org/TR/2011/WD-sparql11-federated-query-20111117/}. 
A recent scientific treatment of the current specification is in
\cite{springerlink:10.1007/978-3-642-21064-8-1}. In this paper, the
authors also show an optimization strategy based on execution order of
so-called well-designed patterns.

For benchmarking federated query processing strategies,
\cite{Schmidt:2011:FBS:2063016.2063054} report the development of
\emph{FedBench}. In this paper, the authors note that the
absence of statistics has very adverse effects on performance as the join
order is impossible to optimize. It is also pointed out that it causes
other problems as well, for example a high number of requests to
federation members. I take this as a key research problem.

This problem has already received some attention, in \cite{5337556} it
is shown how various statistics-gathering techniques, including
histograms, can be computed to assist the SemWIQ optimizer, written by
the same authors.  However, histogram approaches generally suffer
from the problem that they grow too large or become an insufficiently
accurate digest, especially in the face of very heterogeneous data.
\cite{Harth:2010:DSO:1772690.1772733} introduced QTrees, which may
alleviate the problem of histogram size, but which may not solve it.

@@

Numerous benchmarks have been developed for SPARQL, but
\cite{Duan:2011:AOC:1989323.1989340} showed that currently most
benchmarks poorly represent the typical data and queries that are used
on the Semantic Web.
Most recently, \cite{mxro:Morsey2011DBpedia} addressed some of these
problems by using real data and real queries from DBpedia.

>>>>

I take the technology state of the art to be represented by the
current basic SPARQL 1.1 Federated Query Working
Draft\footnote{http://www.w3.org/TR/2011/WD-sparql11-federated-query-20111117/}. In
addition, many have implemented federation that doesn't require 
explicit references to service endpoints, e.g.~\cite{springerlink:10.1007/978-3-642-25073-6-38}.
A recent scientific treatment of the current specification is in
\cite{springerlink:10.1007/978-3-642-21064-8-1}. In this paper, the
authors also show an optimization strategy based on execution order of
so-called well-designed patterns.

A recent review of the state of the art is in
\cite{springerlink:10.1007/978-3-642-17551-0-5}. In addition,
\cite{springerlink:10.1007/978-3-642-25073-6-38} proposes
\emph{bound joins} and proves it can dramatically reduce the number of
requests needed to federation members as well as the implementation of
FedX.

It has been my intention to focus on the two problems listed in
section~3.3.1
in~\cite{springerlink:10.1007/978-3-642-17551-0-5}. Notably, histogram
approaches generally suffer from the problem that they grow too large
or become an insufficiently accurate digest, especially in the face of
very heterogeneous data.  \cite{Harth:2010:DSO:1772690.1772733}
introduced QTrees, which may alleviate the problem of histogram size,
but which may not solve it.

@@

Numerous benchmarks have been developed for SPARQL, but
\cite{Duan:2011:AOC:1989323.1989340} showed that currently most
benchmarks poorly represent the typical data and queries that are used
on the Semantic Web.
Most recently, \cite{mxro:Morsey2011DBpedia} addressed some of these
problems by using real data and real queries from
DBpedia. \cite{Schmidt:2011:FBS:2063016.2063054} has developed a
benchmark for the federated case.
